---
title: "DATA 607 Assignment 7"
author: "Subhalaxmi Rout"
date: "3/15/2020"
output:
  html_document: 
    code_folding: show
    fig_height: 4
    fig_width: 8
    highlight: espresso
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Working with XML and JSON in R

## 1. Introduction

For this assignment, I have created 3 files manually they are:

+ books.html
+ books.xml
+ books.json

I have stored all the file in Github repository and using different library function in R read the file. 

### 1.1 Column names of the data set

There are total 5 columns in the book data set. They are:

+ Name
+ Author
+ Cost
+ Pages
+ PublicationDate

## 2. Load the required libraries

```{r}
#install.packages("XML")
#install.packages("RCurl")
#install.packages("jsonlite")
#install.packages("rvest")
#install.packages("DT")
#install.packages("htmlTable")
library(XML)
library(RCurl)
library(jsonlite)
library(DT)
library(rvest)
library(htmlTable)
```

## 3.HTML File

### 3.1 Code for HTML file 

```{}
<html>

<head>
<title>Book</title>
</head>

<body>

  <table>
  
  <tr>    
    <th>Name</th> 
    <th>Author</th> 
    <th>Cost($)</th> 
    <th>Pages</th> 
    <th>PublicationDate</th>    
  </tr>
  
  <tr>    
  <td>Statistics for Business and Economics</td> 
  <td>James T. McClave, P. George Benson, Terry T Sincich</td> 
  <td>136</td> 
  <td>864</td> 
  <td>12-31-2012</td>    
  </tr>
  
  <tr>    
  <td>OpenIntro Statistics</td> 
  <td>JDavid M Diez, Christopher D Barr, Mine Cetinkaya-Rundel</td> 
  <td>15</td> 
  <td>436</td> 
  <td>07-02-2015</td>    
  </tr>
  
  <tr>    
  <td>An Introduction to Statistical Learning</td> 
  <td>Gareth James, Daniela Witten, Trevor Hastie, Robert Tibshirani</td> 
  <td>52</td> 
  <td>426</td> 
  <td>09-01-2017</td>    
  </tr>
  
  </table>
</body>
</html>
```

### 3.2 Read HTML file from Github repository
```{r}
# get the Github repo URL
htmlURL <- "https://raw.githubusercontent.com/SubhalaxmiRout002/DATA-607-Assignment-7/master/book.html"
#stored the URL
htmlContent <- getURLContent(htmlURL)
#read HTML table
booksHTML <- readHTMLTable(htmlContent, stringsAsFactors=FALSE)
# get all the values
booksHTML <- booksHTML[[1]]
#diaplay table data
datatable(booksHTML)
```

## 4. XML File

### 4.1 Code for XML file

```{}
<records>
  
   <book>
      <Name>Statistics for Business and Economics</Name>
      <Author>James T. McClave, P. George Benson, Terry T Sincich</Author>
      <Cost>136</Cost>
      <Pages>864</Pages>
      <PublicationDate>12-31-2012</PublicationDate>
   </book>
	
   <book>
      <Name>OpenIntro Statistics</Name>
      <Author>JDavid M Diez, Christopher D Barr, Mine Cetinkaya-Rundel</Author>
      <Cost>15</Cost>
      <Pages>436</Pages>
      <PublicationDate>07-02-2015</PublicationDate>
   </book>
   
   <book>
      <Name>An Introduction to Statistical Learning</Name>
      <Author>Gareth James, Daniela Witten, Trevor Hastie, Robert Tibshirani</Author>
      <Cost>52</Cost>
      <Pages>426</Pages>
      <PublicationDate>09-01-2017</PublicationDate>
   </book>
   
</records>
```

### 4.2 Read XML file from Github repository

```{r}
library("methods")
# stored data in a var
xmlURL <- "https://raw.githubusercontent.com/SubhalaxmiRout002/DATA-607-Assignment-7/master/book.xml"
xmlContent <- getURLContent(xmlURL)
#pass the content
booksXMLparsed <- xmlParse(xmlContent)
#put in to data frame
booksXML <- xmlToDataFrame(booksXMLparsed, stringsAsFactors=FALSE)
#view data
datatable(booksXML)
```

## 5.JSON File

### 5.1 Code for JSON file
```{}
{
"book" :[
	{
	"Name" : "Statistics for Business and Economics",
	"Author" : ["James T. McClav", "P. George Benson", "Terry T Sincich"],
	"Cost" : 136,
	"Pages" : 864,
	"PublicationDate" : "12-31-2012"
	},
	{
 	"Name" : "OpenIntro Statistics",
	"Author" : ["JDavid M Diez", "Christopher D Barr", "Mine Cetinkaya-Rundel"],
	"Cost" : 15,
	"Pages" : 436,
	"PublicationDate" : "07-02-2015"
	},
	{
	"Name" : "An Introduction to Statistical Learning",
	"Author" : ["Gareth James", "Daniela Witten", "Trevor Hastie", "Robert Tibshirani"],
	"Cost" : 52,
	"Pages" : 426,
	"PublicationDate" : "09-01-2017"
	}
	]
}
```

### 5.2 Read JSON file from Github repository
```{r}
#stored url in a var
jsonURL <- "https://raw.githubusercontent.com/SubhalaxmiRout002/DATA-607-Assignment-7/master/book.json"
booksJSON <- fromJSON(jsonURL)
# load all the data
booksJSON <- booksJSON[[1]]
#view data
datatable(booksJSON)
```

## 6.Load the information from each of the three sources into separate R data frames

```{r}
str(booksXML)
str(booksHTML)
str(booksJSON)
```

## 7.Compare all 3 data frame

```{r}
identical(booksXML,booksHTML)
```

<span style="color: #354260;background-color: white;">
The `booksXML` and `booksXML` are identical. 
</span>
```{r}
identical(booksXML,booksJSON)

identical(booksHTML,booksJSON)
```

<span style="color: #354260;background-color: white;">
Data set `booksHTML` and `booksJSON` are not identical. 
Dataset `booksJSON` has  `char` for `Name`,`list` data type for `Author`, `numeric` for `Cost`, `numeric` for `Pages` and `char` for `PublicationDate`. However, dataset `booksHTML` has `char` for `Name`,`list` data type for `Author`, `char` for `Cost`, `char` for `Pages` and `char` for `PublicationDate`.
</span>

### 7.1 Try to make all 3 data frame identical
```{r}

# datatype change for cost and pages in HTML
booksHTML$Cost <- as.integer(booksHTML$Cost)
booksHTML$Pages <- as.integer(booksHTML$Pages)

#datatype change for Author in JSON
booksJSON$Author <- as.character(booksJSON$Author)

# datatype change for cost and pages in XML
booksXML$Cost <- as.integer(booksXML$Cost)
booksXML$Pages <- as.integer(booksXML$Pages)

str(booksHTML)
str(booksXML)
str(booksJSON)

identical(booksXML, booksHTML)
identical(booksXML, booksJSON)
identical(booksHTML, booksJSON)
```

Here dataset `booksHTML` and `booksXML` are identical but `booksJSON` is not identical.

## 8.Conclusion

Created all the 3 dataset file on my own. All file has a different format and indentation. R has many functions to read and write these files. This assignment shows, how to read different formated file and get those data. 
We have seen all 3 data frames display the same values for all 3 data sets, but data type is different from one dataset to another. I have tried to change datatype for all 3 data files but in JSON file due to  `Author` type list, not able to get data type character.
From this assignment, got to know more about different file format and how to read and write data from these, this experience will help for data scrapping.